Modelling long-term effects of management strategies in mixed fisheries

### 2019-04-04 doc/MSE_MAR_18.tex
fixed title and intro

\title{MSE and dynamic state variable model with sigma function}

@@

Management strategy evaluation (MSE; \cite{Bunnefeld2011, Sainsbury2000, Smith1994}) seeks to study the implications of management strategies using simulation\cite{Punt2016}. Such simulation should include all important processes of a fishery, which is inherently a socio-ecological system, with coupled dynamics of the fishing fleets, the exploited stocks, and their governance\cite{Punt2016,Rasemeyer2007}. Given the complexity of such a socio-ecological system the uncertainy about the dynamics and feedbacks are consideable. As a result, any forecasted success in achieving goals in a management strategy may thus be inperceivably small compared to the stochasticity in results generated in the system, or worse, be counteracted by unintended consequences of the management strategy. The robustness of management strategies to the uncertainy in the processes that govern fisheries systems thus need to be accounted for \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.

The adaptive response of fishers to their environment is one of the key uncertainties requiring attention in MSE \cite{Fulton2007}. Simulation tools for such adaptive responses are available in fisheries, borrowing methodology from the more general question of state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}. New tools for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018,Batsleer2015, Dowling2011, Poos2010, Gillis1995} have been developed. These models generally predict the effect in the short term (within a fishing trip, or a quota year) by optimizing a utility function and determine which choices best yields the best chance of increasing utility, while keeping track of the state of each individual. The effect of a choice on the utility depends on the economic environment, such as the home port of the vessel and the distance to fishing grounds, and the biological enviroment, such as the spatial distribution of the resources. Using such dynamic variable state models, effects of e.g. relocation costs of marine protected areas have been modelled \cite{Dowling2011}.

Dynamic state variables have also been used to describe discarding behaviour \cite{Batsleer2015, Batsleer2013, Gillis1995}. These studies may shed light on the potential outcomes of the European fisheries management reform of 2013. That reform included using the use of the maximum sustainable yield (MSY) reference points as targets for exploiting commericially important fish stocks and the gradual introduction of a landings obligation (LO) \cite{CFP2013}. Indeed, using spatial and temporal distributions of catch rates within a single year, dynamic state variable models forecasted high costs in mixed fisheries in the short term as a result of the LO. These costs result from removing the option to adapt landings to quota by discarding parts of the catch \cite{Alzorriz2018}. This will result in fishing effort reallocation and early closures of fisheries once quota have been reached. However, the long term benefits to fisheries that could result from improved selectivity are ignored in those studies. Ignoring the potential for improved selectivity leaves out a key element in the perceived benefit of the LO. Because the LO should provide incentives for the use of more selective gears and for fisheries to move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches. Achieving single species MSY while landing all commercial catches in fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Batsleer2013, Ulrich2017}. Forecasting whether such management will be effective depends on understanding the response of individual effort allocation to the management.
This will result in fishing effort reallocation and early closures of fisheries once quota have been reached. However, the long term benefits to fisheries that could result from improved selectivity are ignored in those studies. Ignoring the potential for improved selectivity leaves out a key element in the perceived benefit of the LO. Because the LO should provide incentives for the use of more selective gears and for fisheries to move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches. 
To enable longer term prediction of the effects of changes in fisheries management, and thus complete the MSE, one needs to couple the fleet dynamic models that forecast the fleet response to management to biological dynamic models that forecast the fish population responses to the changing fleet response. We present an MSE that encompasses the individual harvester decision-making process and socioeconomic drivers on management effectiveness, i.e. if the fleet dynamic model results in a short term movement of vessels to allocate effort both spatially and temporally, we also expect longer term changes to occur in fish populations and the economy of the fishery \cite{Alzorriz2018}. Such combination needs to be responsive to all the drivers that influence and motivate the actual fleet and to respect existing constraints \cite{Venables2009}. The drivers will include the (real or perceived) local abundance and catchability of the fished stocks and various resource costs, principally that of fuel. The constraints will include the obvious management-imposed constraints such as levels of total allowable catch, and economical constraints such as the contribution of the annual fines for exceeding catch quotas.

Using a MSE, the main objective of this work is not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependent behaviour of individual fishing vessels. To understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed combined with the LO where discards are not allowed, and to compare it in order to assess on the biological and economic consequences. If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding. Therefore, this manuscript tends to understand the likely adaptive change in fishing patterns that will help us to understand if this leads to a better balance between quotas and catches.

%Since maximizing effort does not maximize catch, the question is if there is an optimum fishing rate that would lead to a maximum yield.

@@

Managament strategy evaluation framework was used to forecast the dynamics of a hypothetical mixed fishery on two species. In the fishery, there are three essential elements: (i) a collection of size strucured fish stocks, whose dynamics are governed by annual reproduction, growth, migration, and mortality; (ii) a management body that evaluates the fishing pressure and aims to set annual quotas in accordance with fishing mortalities dictated by $H_{max}$ reference points (harvest at maximum sustainable yield); and (iii) a fleet of individual fishers who aim to make the best use of their annual quota. Within this mixed fishery, individual vessels make adaptive choices about fishing location and discarding that depend on the distribution of the resources, and the quota that hold. Each of these elements is discussed in more detail below. The essential elements of the framework are summarized in Figure \ref{fig:MSE} and \ref{fig:stockdyn}.

>>>>

\title{Modelling long-term effects of management strategies in mixed fisheries}

@@

Management strategy evaluation (MSE; \cite{Bunnefeld2011, Sainsbury2000, Smith1994}) seeks to study the implications of management strategies using simulations \cite{Punt2016}. Such simulations should include all important processes of a fishery. Any fishery is inherently a socio-ecological system, with coupled dynamics of fishing fleets, exploited stocks, and their governance \cite{Punt2016, Rasemeyer2007}. Given the complexity of socio-ecologicals system the uncertainy about the dynamics and feedbacks are consideable. As a result, any forecasted success in achieving goals of management strategies may thus be small compared to system stochasticity, or worse, be counteracted by unintended consequences of the management strategy. The robustness of management strategies to the uncertainy in the processes that govern fisheries systems thus need to be accounted for \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.

The adaptive response of fishers to their environment is one of the key uncertainties requiring attention in MSE \cite{Fulton2007}. Simulation tools for such a response are available, borrowing methodology from state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}. State-dependent behaviour of individual fishing vessels, aggregated into fleet dynamics, have been developed using stochastic dynamic programming \cite{Alzorriz2018,Batsleer2015, Dowling2011, Poos2010, Gillis1995}. These models generally predict short term (within a fishing trip, or a quota year) effects of changes in management or the environment by determining which choices yields the best chance of increasing utility, while keeping track of the state of each individual. The effect of a choice on utility depends on the state of the individual, the economic environment, such as fuel prices, and the biological enviroment, such as the spatial distribution of resources.

Using dynamic variable state models, effects of e.g. relocation costs of marine protected areas have been modelled \cite{Dowling2011}. Dynamic state variables have also been used to describe discarding behaviour \cite{Batsleer2015, Batsleer2013, Gillis1995, Alzorriz2018}. Some of these studies have tried to shed light on the potential outcomes of the European fisheries management reform of 2013. That reform included using the use of the maximum sustainable yield (MSY) reference points as targets for exploiting commercially important fish stocks and the gradual introduction of a landings obligation (LO) \cite{CFP2013}. Indeed, using spatial and temporal distributions of catch rates within a single year, dynamic state variable models forecasted high costs in mixed fisheries in the short term as a result of the LO \cite{Alzorriz2018, Batsleer2013, Batsleer2015}. These costs result from removing the choice to discard part of the catch and in doing so adapt landings to quota \cite{Alzorriz2018}. This will result in fishing effort reallocation and early closures of fisheries once quota have been reached.

Meanwhile, the LO potentially improves selectivity of fisheries by providing incentives to move away from areas with high levels of unwanted catch and the use of more selective gears \cite{Alzorriz2016, Alzorriz2018}. The improved selectivity should lead to higher long term catches. These long term benefits to fisheries that could result from improved selectivity under the LO are ignored in previous studies on the adaptive response of fisheries \cite{Batsleer2015, Batsleer2013, Alzorriz2018}. Ignoring the potential for improved selectivity leaves out a key element in the perceived benefit of the LO. In mixed fisheries, however, optimizing single species selectivity and achieving MSY while landing all commercial catches is challenging because of trade-offs in the catches of different species and size-classes \cite{Batsleer2013, Ulrich2017}. Forecasting whether any management will be effective requires understanding the response of fishers constrained by these trade-offs and the effects of the response on the development of fish populations \cite{Venables2009}. If the fleet dynamic model results in a short term movement of vessels to allocate effort both spatially and temporally, we also expect longer term changes to occur in fish populations and the economy of the fishery \cite{Alzorriz2018}.

We present an MSE framework that couples forecasts from a state-dependent behaviour model to a biological dynamic models that forecasts population effects resulting from the collective fishers behaviour. The model includes the (real or perceived) local biomass of two populations, governed by growth, mortality, and migration. These populations are size-structured, and are modelled in a spatially heterogeneous environment. The populations are harvested by a single fleet that has annual fishing rights for both populations, consisting of individuals who make monthly choices on fishing locations. These choices depend on the perceived spatial distributions of biomasses, but also on various costs, principally that of fuel. The constraints on the fleet  includes management-imposed constraints such as fish quotas. In the framework, these quotas are decided by a a management body that observes the total biomass of the two stocks and with the goal of constraining fishing mortality so that maximum sustainable yields are achieved.

The MSE framework is usedt o understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed combined with the LO where discards are not allowed, and to compare it in order to assess on the biological and economic consequences. If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding. Therefore, this manuscript tends to understand the likely adaptive change in fishing patterns that will help us to understand if this leads to a better balance between quotas and catches.

@@

The managament strategy evaluation framework was used to forecast the dynamics of a hypothetical mixed fishery on two species. In the fishery, there are three essential elements: (i) a collection of size strucured fish stocks, whose dynamics are governed by annual reproduction, growth, migration, and mortality; (ii) a management body that evaluates the fishing pressure and aims to set annual quotas in accordance with fishing mortalities dictated by $H_{max}$ reference points (harvest at maximum sustainable yield); and (iii) a fleet of individual fishers who aim to make the best use of their annual quota. Within this mixed fishery, individual vessels make adaptive choices about fishing location and discarding that depend on the distribution of the resources, and the quota that hold. Each of these elements is discussed in more detail below. The essential elements of the framework are summarized in Figure \ref{fig:MSE} and \ref{fig:stockdyn}.
